---
title: "Literate Analysis: Ocean Water Samples"
author: "Angie Kochkarova"
format: html
execute:
  echo: true
---

# Introduction

This document represents a literate analysis practice using seawater chemistry data. The data examined is the nitrogen and phosphate concentrations in arctic oceans to explore how closely the samples match the Redfield ratio (16:1), which reflects nutrient balance in marine ecosystems.

# About the data

The dataset **BGchem2008data.csv** was obtained from the Arctic Data Center (https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A35ad7624-b159-4e29-a700-0c0770419941). 

The data was downloaded on August 8, 2025.

# Setup

```{r}
library(readr)
library(here)
```

```{r}
## Read in Data

#| message: false
#| warning: false

bg_chem <- read_csv(here::here("data","raw", "BGchem2008data.csv"))
```
```{r}
#| eval: false
#| echo: false

colnames(bg_chem)
str(bg_chem)
head(bg_chem)
summary(bg_chem)
unique(bg_chem$Date)
```

# Analysis

We calculate the mean nitrate (NO3), nitrite (NO2), ammonium (NH4), 
and phosphate (P) concentrations using the mean() function.

```{r}
```{r}
nitrate <- mean(bg_chem$NO3, na.rm = TRUE)
nitrite <- mean(bg_chem$NO2, na.rm = TRUE)
amm     <- mean(bg_chem$NH4, na.rm = TRUE)
phos    <- mean(bg_chem$P,   na.rm = TRUE)
```

## Ratio of Nitrogen vs Phosphorous  
```{r}
```{r}
ratio <- (nitrate + nitrite + amm) / phos
ratio
```

The Redfield ratio for this dataset is approximately `r round(ratio, 2)`.

# Plot RedField Ratio

```{r}
#| echo: false

plot(
  bg_chem$P,
  bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4,
  xlab = "Phosphate (P)",
  ylab = "Total Nitrogen (NO2 + NO3 + NH4)",
  main = "Ocean Water Samples: Total Nitrogen vs Phosphate"
)
```

# Conclusion

About the results:

Observing that the Nitrogen (N) to Phosphorous (P) ratio is 6.22, this is much lower than RedField ratio of 16. From the generated scatterplot there shows a positive correlation between N and P, which is ideal. Though, what makes the ratio lower than the ideal RedField ratio is the amount of N present. This indicates that the arctic samples are relatively nitrogen-limited, and therefore will not be ideal for sustaining diverse life forms.

In this mini-analysis, we:

Loaded packages in a dedicated setup chunk

Read in BGchem2008data.csv using a project-root-relative path (here())

Calculated simple nutrient means

Computed a mean nitrogen:phosphate ratio

Created a basic plot relating total nitrogen to phosphate



