---
title: "Cleaning and Wrangling Data with Alaskan Salmon"
author: "Angie Kochkarova"
---

# Introduction

Alaska salmon fisheries are an important ecological and economic resource, both globally and locally. Salmon fishing supports commercial fishing industries, providing jobs in coastal communities, and management of fisheries play a key role in marine and freshwater ecosystems. Because salmon migrate across large geographic areas, the catch levels can change depending on environmental conditions, regional productivity, and management strategies.

By analyzing differences in average salmon catch across management regions this will help researchers reveal patterns in regional productivity, the pressures of fishing on the ecosystem, and population counts. Through understanding these differences improvements can be made to resource management, sustainability planning, and long-term conservation of salmon populations.

# Data

```{r}
# Load required packages
library(dplyr)
library(tidyr)
library(readr)
#| message: false
```

```{r}
# Import dataset from online repository
catch_original <-read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

## Data Sources

Data were obtained from the KNB Data Repository: Alaska commercial salmon catches by management region (1886–1997). File used: df35b.302.1

URL: https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1

Data downloaded March 2nd, 2026.

# Analysis

```{r}
mean_region <- catch_original %>% # Start with original dataset
  select(-All, -notesRegCode) %>% # Remove unnecessary columns
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>%  # Replace "I" with "1" in Chinook column
  mutate(Chinook = as.numeric(Chinook)) %>% # Convert Chinook column to numeric values
  pivot_longer(-c(Region, Year), # change data format form wide to long
               names_to = "species",# Create new column for species names
               values_to = "catch") %>% # Create new column for catch values
  mutate(catch = catch * 1000) %>% # Convert catch values to full numbers
  group_by(Region) %>%  # Group the data by region
  summarize(mean_catch = mean(catch)) %>% # Calculate the mean catch per region
  arrange(desc(mean_catch)) # Sort from highest to lowest mean catch

head(mean_region)
```

## Discussion

From the analysis, we can determine that the region of SSE has the highest mean of salmon catches in comparison to the other regions. After SSE, the next region with the highest production of salmon is BRB. The differences of salmon catch and production per region indicate that there are probable differences in migration style and how the fisheries are managed. 
