---
title: "a05 Assignment"
author: "Angie Kochkarova"
Date: 04-02-2026
---

## Introduction

Through the MetaData collected from the Sacramento-San Joaquin Delta, we as researchers are able to gain a better understanding of how people interact with local ecological restoration areas. This work is important as it allows us to evaluate the impact humans make on these areas and use the resources provided by the area. In this analysis, we specifically explore visitation patterns across restoration locations in the Sacramento-San Joaquin Delta. Through the visualization the number and type of visitors through ggplot2, we can better understand human activity and impact within these areas.

## Data

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(forcats) # makes working with factors easier
library(ggplot2)
library(leaflet) # interactive maps
library(DT) # interactive tables
library(scales) # scale functions for visualization
library(janitor) # expedite cleaning and exploring data
library(viridis) # colorblind friendly color palette
```

```{r}
install.packages("leaflet")
install.packages("DT")
install.packages("janitor")
install.packages("viridis")
```

```{r}
delta_visits_raw <- read_csv(here::here("data", "raw","Socioecological_monitoring_data.csv"))
```

```{r}
## Check out column names

colnames(delta_visits_raw)

## Peak at each column and class
glimpse(delta_visits_raw)

## From when to when
range(delta_visits_raw$Date)

## Which time of day?
unique(delta_visits_raw$Time_of_Day)
```

```{r}
# Clean column names to snake_case
delta_visits <- delta_visits_raw %>%
  clean_names()

# Convert data from wide to long format
visits_long <- delta_visits %>%
  pivot_longer(
    cols = c(sm_boat, med_boat, lrg_boat, bank_angler, scientist, cars),
    names_to = "visitor_type",      # new column for visitor categories
    values_to = "quantity"          # values stored here
  ) %>%
  rename(restore_loc = eco_restore_approximate_location) %>%  # rename column
  select(-notes)   # remove unnecessary column
```

```{r}
# Calculate total daily visits grouped by location, date, and visitor type
daily_visits_loc <- visits_long %>%
  group_by(restore_loc, date, visitor_type) %>%  # group data
  summarise(daily_visits = sum(quantity), .groups = "drop")  # sum visits
```

###Plot of Data: Total Visits by Location and Visitor Type

```{r}
# Create a bar plot showing total visits by location and visitor type
ggplot(data = daily_visits_loc,
       aes(x = daily_visits, y = restore_loc, fill = visitor_type)) +  # map variables
  geom_col() +  # create bar chart
  scale_fill_viridis_d() +  # apply colorblind-friendly colors
  labs(
    x = "Number of Visits",
    y = "Restoration Location",
    fill = "Visitor Type",
    title = "Total Visits by Location and Visitor Type",
    subtitle = "Sacramento-San Joaquin Delta Monitoring Data"
  ) +
  theme_minimal()  # clean theme
```

### Interactive Map with Unique Locationa and Coordinates

```{r}
# Create dataset of unique locations with coordinates
locations <- visits_long %>%
  distinct(restore_loc, .keep_all = TRUE) %>%  # keep unique locations
  select(restore_loc, latitude, longitude)     # select relevant columns

# Create interactive map
leaflet(locations) %>%
  addTiles() %>%  # add base map
  addMarkers(
    lng = ~longitude,   # longitude for markers
    lat = ~latitude,    # latitude for markers
    popup = ~restore_loc  # label popup
  )
```

## Discussion

Two Visuals were created to interpret the Socioecological monitoring data from the Sacramento-San Joaquin Delta, a plot and an interactive map.

The plot, specifically a bar plot, shows the variation in visitation across different restoration locations and highlights the different visitor types. Depending on the location, some areas receive higher traffic from boaters, while others show more activity from anglers or scientists. This shows that the type of visitation depends on the area and ecology of the area.

As for the interactive map, it's a great visual to show the distance of these areas from each other, and gives the visual context as to which areas have been restored. These clusters of locations indicate how important those ecological areas are to the community and the protection/saftey of the local species.

Together, these two visualizations are great way to monitor useage and visitation of these areas by the community, and provide useful information to scientists and parks and rec. rangers for future management and conservation efforts.
