---
title: "R Practice: Tidy Data and Joins"
author: "Angie Kochkarova"
---

```{r}
library(readr)
library(dplyr)
```

```{r}
library(here)
library(lubridate) # for bonus question
```

## Quick question: Did you get a message after loading the libraries? What is it telling you? Talk to your neighbor about it or write a note in your qmd.

### Yes, I got the message that the package 'lubridate' was attached, and that some objects within the base package were being masked. The objects masked were: date, intersect, setdiff, and union. The masked objects come from the "base" package, or what R already provides. Because both R and lubridate have these 4 functions/objects, yet lubridate was loaded in after R, the functions from lubridate are prioritized. Therefore the functions from lubridate become default in this session.

## read in data

```{r}
bird_observations <- read_csv(here::here("data","raw","52_pp52_birds_1.csv"))

surveys <- read_csv(here::here("data", "raw","52_pp52_surveys_1.csv"))

sites <- read_csv(here::here("data","raw","52_pp52_sites_1.csv"))

taxalist <- read_csv(here::here("data","raw","52_pp52_taxalist_1.csv"))
```

```{r}
View(surveys)
# returns dimensions of the dataframe by number of rows and number of cols
dim(bird_observations)
```

```{r}
# returns the top six rows of the dataframe
head(bird_observations)
```

```{r}
# returns all the columns and some info about the cols
glimpse(bird_observations)
```
```{r}
# similar to glimpse but returns some summary statistics about the cols
summary(bird_observations)
```
```{r}
# returns column names 
names(bird_observations)
```
```{r}
# returns unique values in a column. In this case we can see all the different bird species IDs
unique(bird_observations$species_id)
```

```{r}
birds_subset <- bird_observations %>% # create subset of codes for bird observations
  filter(species_id %in% c("BHCO", "RWBL")) %>% # only want birds with species ID BHCO and RWBL
  filter(site_id %in% c("LI-W", "NU-C")) # only want from sites LI-W and NU-C
```

```{r}
# syntax using pipe
birds_left <- birds_subset %>% 
  left_join(y = sites, by = "site_id")
# don't see x = birds_subset here because piping in birds_subset means it automatically assumes birds_subset as x.

# syntax without pipe
birds_left <- left_join(x = birds_subset, y = sites, by = "site_id")
head(birds_left)
```

```{r}
birds_full <- full_join(x = birds_subset, y = sites, by = "site_id")
birds_inner <- inner_join(x = birds_subset, y = taxalist, by = "species_id" )
head(birds_inner)
```

```{r}
names(bird_observations)
```

```{r}
bird_obs_subset <- bird_observations %>% 
  full_join(y = taxalist, by = "species_id") %>% 
  full_join(y = sites, by = "site_id") %>% 
  full_join(y = surveys, by = c("site_id", "survey_id")) %>%  
  select(survey_date, common_name, park_name, bird_count, observer)
```

```{r}
unique(bird_obs_subset$observer)
```
```{r}
bird_obs_subset <- bird_obs_subset %>% 
  mutate(observer = if_else(condition = observer == "Josh Burns", 
                            true = "J. Burns", 
                            false = observer))

unique(bird_obs_subset$observer)
```

